Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2020-01-25 manuscript/cover-letter/cover_letter.tex
Revise cover letter
Reduce length and directly address four primary questions asked by the the
journal.

\begin{newlfm}
  Please find attached our manuscript entitled ``Integrating genotypes and phenotypes improves long-term forecasts of seasonal influenza A/H3N2 evolution''.
  We would be grateful if you considered it for publication in \textit{eLife}.
  We believe that \textit{eLife} is an appropriate venue, as this work provides the most comprehensive analysis of fitness metrics for seasonal influenza to date and a generalized forecasting framework that will be a valuable tool for other biological researchers.

  The selection of vaccine strains for seasonal influenza A/H3N2 remains a daunting challenge due to constant accumulation of mutations in the hemagglutinin (HA) surface protein.
  Until recently, experts selected vaccine strains by manually synthesizing experimental measurements of antigenic drift provided by hemagglutination inhibition (HI) assays.
  Two landmark papers in 2014 showed that H3N2 evolution may be predictable from HA sequences alone.
  The first paper, ``A predictive fitness model for influenza'' by {\L}uksza and L\"assig in \textit{Nature}, established a mechanistic fitness model that predicted the frequencies of phylogenetic clades one year in advance.
  This model defined three biological fitness metrics including antigenic drift measured by mutations at epitope sites, functional constraint measured by mutations at non-epitope sites, and recent clade growth measured by synonymous mutations.
  The second paper, ``Predicting evolution from the shape of genealogical trees'' by Neher et al. in \textit{eLife}, developed a phylogenetic metric of recent clade fitness called the local branching index (LBI).
  This metric accurately identified high fitness clades without providing a biological basis for that success.
  Predictions from both of these models are now routinely considered in the World Health Organization's vaccine composition meetings.

  Several substantial questions remained unanswered by these models.
  Are genetic estimates of antigenic drift sufficient replacements for the gold standard phenotypic measures provided by HI assays?
  What mechanistic components of influenza fitness does LBI measure?
  Can simple genetic estimates of functional constraint be improved upon using modern deep mutational scanning (DMS) assays that measure the functional effects of every single amino acid mutation to a given strain's HA?
  Here, we investigate for the first time the relationships between all of these genotypic and phenotypic estimates of influenza fitness using a collection of HA sequences and HI measurements of an unprecedented breadth and depth.
  To this end, we developed a novel approach to forecasting influenza evolution by estimating the sequence composition of future populations and rigorously tested this framework with simulated and natural H3N2 populations.
  We implemented new experimentally-informed fitness metrics including antigenic drift by HI assays and functional constraint by DMS assays.
  We applied modern machine learning methods for model training, validation, and testing to thoroughly evaluate the predictive accuracy of these new metrics as well as {\L}uksza and L\"assig's epitope and non-epitope metrics and Neher et al.'s LBI.
  Importantly, we verified that this framework can both capture dynamics of phylogenetic clades and identify optimal candidate strains for vaccine composition.

  We find that phenotypic measures of antigenic drift consistently outperform genetic estimates based on epitope sites.
  We further show that the predictive value of these epitope sites in {\L}uksza and L\"assig's mechanistic model was most likely due to inadvertantly borrowing information from the future during model fitting.
  This surprising finding has important ramifications for future attempts to map influenza genotypes to fitness, the use of epitope sites in practical forecasts, and the continued value of HI assays.
  Additionally, we find that DMS-based estimates of functional constraint produced models that overfit to the genetic background of the strain, A/Perth/16/2009, used in the experiments.
  Instead, we find that {\L}uksza and L\"assig's original minimal model of mutational load at non-epitope sites is more robust.
  Composite models that integrate mutational load with either HI phenotypes or LBI performed the best and indicate that LBI may be measuring similar fitness effects as HI assays.
  We have integrated these models into nextstrain.org, enabling real-time forecasts for H3N2 for influenza researchers, decision makers, and the public.

  Our findings challenge the widely held assumption about the power of epitope sites to predict the success of influenza populations.
  Correspondingly, we validate the predictive value of HI assays to measure antigenic drift and provide a framework that synthesizes these measurements into forecasts for expert interpretation.
  Beyond influenza, our framework enables similar forecasting efforts for other viral and bacterial pathogens by allowing researchers to define sample-specific fitness metrics in tidy data frames and fit models to those metrics without relying on phylogenetic inference.
  We are confident this open source framework will provide a foundation on which future forecasting efforts can be built or inspired.
  For these reasons, our software and findings are relevant not only to the specific audience of influenza researchers, but also to the more general audience of infectious disease researchers, evolutionary biologists, and bioinformaticians.
\end{newlfm}

>>>>

\begin{newlfm}
  Please find attached our manuscript entitled ``Integrating genotypes and phenotypes improves long-term forecasts of seasonal influenza A/H3N2 evolution''.
  We would be grateful if you considered it for publication in \textit{eLife}.

  The World Health Organization selects vaccine strains for seasonal influenza A/H3N2 based on experimental measurements of antigenic drift and computational models of hemagglutinin (HA) sequence evolution.
  Hemagglutination inhibition (HI) assays provide phenotypic measurements of antigenic drift that have been the gold standard for decades.
  In their 2014 \textit{Nature} paper, {\L}uksza and L\"assig provided the first mechanistic fitness model to predict the success of H3N2 populations one year into the future from HA sequences alone.
  That same year in \textit{eLife}, Neher et al. presented the local branching index (LBI), a fitness estimate based on rapid branching in HA phylogenies.
  No studies to date have integrated these genetic and phenotypic measures of HA into a single model to understand their relative contributions to influenza fitness.

  To address this gap, we present the first fitness framework to integrate genetic and phenotypic measures of HA antigenic drift and functional constraint and phylogenetic measures of clade growth.
  We implemented novel measures of antigenic drift by HI assays, functional constraint measured by deep mutational scanning (DMS) experiments, and clade growth based on recent changes in clade frequencies.
  We also reimplemented {\L}uksza and L\"assig's sequence-based measures of antigenic drift and functional constraint, and Neher et al.'s clade growth metric, LBI.
  To identify the most predictive models, we tested each of these fitness metrics individually and in relevant combinations using time-series cross-validation.
  The most accurate models included a combination of sequence-based functional constraint and either HI-based measures of antigenic drift or LBI-based measures of clade growth.

  Our findings answer several open questions in the field and point the way for future forecasting efforts.
  We find that genetic estimates of antigenic drift are insufficient replacements for the gold standard phenotypic measures provided by HI assays.
  We further show that the success of previous models based on these genetic estimates is due, in part, to inadvertantly borrowing information from the future.
  In contrast, we show that simple genetic estimates of functional constraint are superior to more detailed estimates provided by deep mutational scanning (DMS) measurements from a single background strain.
  The results suggest that future forecasting models should strive not to replace genetic fitness metrics with phenotypic metrics but to identify metrics that are the least likely to be affected by historical contingency.

  We designed our fitness framework as a reproducible workflow built from open source tools, released as open source itself, and integrated into the real-time pathogen surveillance ecosystem of Nextstrain.
  Our models estimate the sequence composition of future populations instead of phylogenetic clade frequencies, enabling subsequent forecasting efforts for rapidly evolving organisms whose genomes are not suited for phylogenetic inference such as recombining viruses and bacteria.
  Importantly, fitness metrics can be readily defined in tidy data frames from any continuous trait of a sample.
  Altogether, this manuscript and the associated fitness framework are immediately relevant to a general audience of evolutionary biologists, infectious disease researchers, virologists, and bioinformaticians.
  For these reasons, we believe that \textit{eLife} is an appropriate venue for this work.
\end{newlfm}
